---
title: "Data Analysis on Neigbourhoods in Toronto Based on Bicycle Theft Cases"
author: "Pande Adhi Prathama"
date: "01/10/2021"
output:
  html_document:
    df_print: paged
---

# Prologue

In this exercise, we aim to perform cluster analysis of neighbourhoods in Toronto based on bicycle theft cases which reported from 2014 - 2020.

To perform the analysis, we use following libraries.

```{r, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(factoextra)
library(ggpubr)
library(knitr)
library(janitor)
```

The data set we use to perform the analysis is provided by Toronto Police Service (link: <https://data.torontopolice.on.ca/datasets/bicycle-thefts/explore?location=43.727864%2C-79.122754%2C9.64>).

Some variables from original dataset have been removed from the analysis, with details as follow:

-   "X", "Y", "Longitude", and "Latitude": as we are not concerned with the exact geographical detail of each theft case.

-   "Hood_ID": as this variable contains similar information to "NeighbourhoodName".

-   Details of Occurrence and Report Dates (e.g Day, Month, Year, etc), as we can refer to "Occurrence Date" and "Report Date".

-   Detailed information of stolen bicycles, apart from 'Bike_Type' variable: as we are more interested to the number of case instead of bike.

-   "OBJECTID", "event_unique_id", and "ObjectId2": as these variables only show up 

```{r, message=FALSE}
bike_theft <- read.csv(url("https://raw.githubusercontent.com/adhiprathama92/toronto_bicycle_theft_data/main/Bicycle_Thefts.csv"))
bike_theft <- bike_theft %>%
  select(-c("ï..X", "Y", "Hood_ID", "Occurrence_Year", "Occurrence_Month","Occurrence_DayOfWeek","Occurrence_DayOfMonth","Occurrence_DayOfYear","Occurrence_Hour","Report_Year", "Report_Month","Report_DayOfWeek","Report_DayOfMonth","Report_DayOfYear","Report_Hour", "Longitude", "Latitude", "Bike_Make", "Bike_Model", "Bike_Speed", "Bike_Colour", "Cost_of_Bike","OBJECTID", "event_unique_id", "ObjectId2"))
bike_theft
```

# Understanding Data

We notice that majority of variables have 'character' data type.

```{r}
str(bike_theft)
```

'Occurrence_Date' and 'Report_Date' variables are classified as 'character' instead of 'date and time'. Hence, we need to reclassify those variables to 'POSIXct'.

Interestingly, we also notice that there are 439 entries which shows gap year between Occurrence Date and Report Date. Up until now, there is no information found on why this is happened. From the time being, we will utilise **Report Date** for our analysis as we assume the police will do investigation right after the case is reported.

```{r}
#change occurance date and report date tp POSIXct data type
bike_theft <- bike_theft %>%
  mutate(Occurrence_Date = as.POSIXct(Occurrence_Date), Report_Date = as.POSIXct(Report_Date))

str(bike_theft)
```

```{r, message=FALSE}
#figure out gap year between occurrence vs report
bike_theft %>%
  mutate(Occurrence_Year = year(Occurrence_Date), Report_Year = year(Report_Date)) %>%
  group_by(Occurrence_Year, Report_Year) %>%
  summarise(Gap_Occurrence_Report = Report_Year - Occurrence_Year, frequency = n()) %>%
  distinct() %>%
  filter(Gap_Occurrence_Report > 0)
```

Further, on 'Bike_Type' variable, the data author also provide definition of each abbreviation as follow. (refer to: <https://torontops.maps.arcgis.com/sharing/rest/content/items/332f6e958b704d9aa023e7c3487f710f/data>)

-   BM : BMX
-   EL : Electric
-   FO : Folding
-   MT : Mountain
-   RC : Racer
-   RE : Recumbant
-   RG : Regular
-   SC : Scooter
-   TA : Tandem
-   TO : Touring
-   TR : Tricycle
-   UN : Unicycle
-   OT : Other
-   UNKNOWN : Type Unknown

```{r}
unique(bike_theft$Bike_Type)
```

# Data Cleaning

## Figure out missing data

By running the following code, we notice that there are no variables with missing data

```{r}
bike_theft %>%
  sapply(function(x) sum(is.na(x)))
```

## Figure out variables which have zero (or near zero) variance

In this section, we are interested to know which variables that have zero or near zero variance. This is helpful to make our data clean from variables which has (1) no unique values or (2) significant modus value outnumbered others.

```{r}
caret::nearZeroVar(bike_theft,saveMetrics = TRUE)
```
<br>
Notice that though **'Status'** and **'City'** variables still has unique values (zeroVar = FALSE), but their respective modus value is obviously outnumbered the second modus by looking at freqRatio. Hence, their near zero variance (nzv) is classified as TRUE.

The following barchart shows up all available value in **"Status"** variable. We notice that 'Stolen' value is far outnumbered 'Unknown' and 'Recovered', meaning that we might not find any insight from this variable.

Hence, we can delete **"Status"** variable.

```{r, fig.width=12}
bike_theft %>%
  count(Status) %>%
  ggplot(aes(y = reorder(Status, n), x = n, fill = reorder(Status, n))) + 
  geom_bar(stat="identity") + 
  theme_bw() + 
  geom_text(aes(label=n), position=position_dodge(width=0.8), hjust= -0.2) +
  xlab("Number of Cases") + 
  ylab("Status") +
  scale_fill_brewer(palette="Dark2") +
  ggtitle("Number of Cases by Case Status") +
  theme(
    legend.title = element_blank(),
  )
```

<br>For **'City'** variable, we notice that there is 'NSA' value. 'NSA' stands for 'Not Specific Area', meaning the theft case is either located outside the city of Toronto or invalidated locations ( refer to: <https://torontops.maps.arcgis.com/sharing/rest/content/items/c0b17f1888544078bf650f3b8b04d35d/data> ).

As such, we will take out also 'City' variable.

```{r, fig.width=12}
bike_theft %>%
  count(City) %>%
  ggplot(aes(y = reorder(City, n), x = n, fill = reorder(City, n))) + 
  geom_bar(stat="identity") + 
  theme_bw() + 
  geom_text(aes(label=n), position=position_dodge(width=0.8), hjust= -0.2) +
  xlim(0, 30000) +
  xlab("Number of Cases") + 
  ylab("City") +
  ggtitle("Number of Cases by City") +
  scale_fill_brewer(palette="Dark1") +
  theme(
    legend.title = element_blank(),
  )
```

## Cleaning "NeighbourhoodName" Data

We also notice there is NSA value in "NeighbourhoodName" variable, hence the entries with "NSA" value in "NeighbourhoodName" variable are also taken out.


```{r}
bike_theft %>%
  count(NeighbourhoodName) %>%
  arrange(n)
```

## Cleaned Dataset

Here is our dataset looks like after taking out 'Status', 'City', and 'Occurrence_Date' variables, in addition to remove entries with 'NSA' value in 'NeighbourhoodName' variable.

```{r}
bike_theft <- bike_theft %>%
  select(-c(Occurrence_Date, City, Status)) %>%
  filter(NeighbourhoodName != "NSA")
bike_theft
```

# Analysis

## Exploring Cases by Neighbourhood

To explore cases by neighbourhood, we are interested to utilise Pareto analysis to determine which neighbourhood that are:
- being in top 10 contributors to overall bicycle theft cases
- outside top 10 but still contribute to 80% of all bicycle theft cases
- rest of neighbourhoods that do not fall into above categories

To do this, we calculate number of cases by Neighbourhood and arrange in descending manner. Here we add "Cummulative_Percentage" variable to indicate cummulative contribution of each neighbourhood.

```{r}
bike_theft_by_neighbourhood <- bike_theft %>%
  group_by(NeighbourhoodName) %>%
  summarise(Frequency = n()) %>%
  arrange(desc(Frequency)) %>%
  mutate(Cummulative_Percentage = cumsum(Frequency)*100/nrow(bike_theft))
bike_theft_by_neighbourhood
```

<br>
From here, we can see **10 neighbourhoods racks up around 47% of overall bike theft cases** in Toronto, while **other 32 neighbourhoods add up the number to around 80%**. Meanwhile the rest **98 neighbourhoods only contributes to 20% of overall bike theft cases**.

We classify neighbourhoods into 3 groups called **"Top 10 Neighbourhoods"**, **"Next Top 32 Neighbourhoods"**, **"Rest of Neighbourhoods"** going forward.

```{r}
# List of Top 10 Neighbourhoods (contribution to overall theft cases = 47%)
top_10_neighbourhood <- bike_theft %>%
  group_by(NeighbourhoodName) %>%
  summarise(Number_of_Cases = n()) %>%
  arrange(desc(Number_of_Cases)) %>%
  head(10)

# List of Next Top 32 Neighbourhoods (contribution to overall theft cases = 33%)
next_top_32_neighbourhood <- bike_theft %>%
  filter(!NeighbourhoodName %in% c(unique(top_10_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(32)

# List of The Rest 98 Neighbourhoods (contribution to overall theft cases = 20%)
other_neighbourhood <- bike_theft %>%
  filter(!NeighbourhoodName %in% c(unique(top_10_neighbourhood$NeighbourhoodName)) & 
          !NeighbourhoodName %in% c(unique(next_top_32_neighbourhood$NeighbourhoodName))
         ) %>%
  group_by(NeighbourhoodName) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

Here are the list of Top 10 Neighbourhoods, including their respective number of cases.

```{r}
top_10_neighbourhood %>%
  ggplot(aes(x=Number_of_Cases, y=reorder(NeighbourhoodName, Number_of_Cases))) +
  geom_bar(stat="identity", fill = "#FFDB6D") +
  theme_bw() + 
  geom_text(aes(label=Number_of_Cases), size = 3, position=position_dodge(width=0.8), hjust= -0.2) +
  xlim(0, 3000) +
  xlab("Number of Cases") + 
  ylab("Neighbourhood") +
  ggtitle("Top 10 Neighbourhoods by Number of Bicycle Theft Cases")+
  theme(
    plot.title = element_text(size = 10, face = "bold", hjust = -2.2),
    axis.title = element_text(size = 8, face = "bold"),
    axis.text = element_text(size = 8)
  )
```


## Exploring Cases by Premise Type

```{r, message=FALSE}
# bar chart of number of cases across neighbourhoods by premise type
premise_type_overall <- bike_theft %>%
  group_by(Premises_Type) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x=reorder(Premises_Type, -n), y= n)) + 
  geom_bar(stat='identity', fill = "#2B60DE") +
  geom_text(aes(label=n), position=position_dodge(width=0.5), vjust= -0.5, size = 3) +
  ylab("Number of Cases") + 
  xlab("Premise Type") +
  ggtitle("Overall Neighbourhoods") +
  theme_bw()+
  ylim(0, 9000)+
  theme(
    legend.title = element_blank(),
    plot.title = element_text(size = 10, hjust = -0.095, face ="bold"),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 9)
  )

# boxplot of number of cases in Top 10 neighbourhoods by premise type
premise_type_top_10 <- bike_theft %>%
  filter(NeighbourhoodName %in% c(unique(top_10_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, Premises_Type) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(Premises_Type, levels=c("Outside", "Apartment", "House", "Commercial", "Other", "Educational", "Transit")), y = count)) + 
  geom_boxplot(fill = "#FFDB6D", color = "#C4961A")+
  xlab("Premise Type") +
  ylab("Number of Cases") +
  ggtitle("Top 10") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# boxplot of number of cases in Next 32 neighbourhoods by premise type   
premise_type_next_32 <- bike_theft %>%
  filter(NeighbourhoodName %in% c(unique(next_top_32_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, Premises_Type) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(Premises_Type, levels=c("Outside", "Apartment", "House", "Commercial", "Other", "Educational", "Transit")), y = count)) + 
  geom_boxplot(color = "#306754", fill = "#50C878") +
  xlab("Premise Type") +
  #ylab("Number of Cases") +
  ggtitle("Next Top 32") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# boxplot of number of cases in Rest 98 neighbourhoods by premise type 
premise_type_others <- bike_theft %>%
  filter(NeighbourhoodName %in% c(unique(other_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, Premises_Type) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(Premises_Type, levels=c("Outside", "Apartment", "House", "Commercial", "Other", "Educational", "Transit")), y = count)) + 
  geom_boxplot(color = "#F70D1A", fill = "#FF69B4") +
  xlab("Premise Type") +
  #ylab("Number of Cases") +
  ggtitle("Rest of Neighbourhoods") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

#combine altogether into a plot
annotate_figure(ggarrange(premise_type_overall,
                          ggarrange(premise_type_top_10,
                                    premise_type_next_32, 
                                    premise_type_others, 
                                    nrow = 1, 
                                    ncol = 3),
                          nrow = 2,
                          ncol = 1),
                top = text_grob(expression(atop(bold("Number of Bicycle Theft Cases in Toronto by Premise Type (2014 - 2020)"), scriptstyle("Top 10 Neighbourhoods represent 47% of total cases, while Next Top 32 and The Rest contribute to 33% and 20% respectively"))))
                )
```
<br>
It is understood that the bicycle parked outside building premises seems more prone to be thieved, followed by residential areas (i.e apartments and houses). This is aligned with cases observed in Top 10 neighbourhoods,

Meanwhile, in Next Top 32 Neighbourhoods and Rest of Neighbourhoods groups, number of cases outside, apartment and house show almost similar median value. 

We might consider premise type as the variable to distingusih "Top 10" among other neighborhoods. To re-confirm this, we can use K-Means Clustering.

## K-Means Clustering

### Data Pre-processing

We only consider 'Neighbourhood' and 'Premise Type' variables as basis of clustering analysis.

Before performing cluster analysis, we need to transform our data so that it displays count of premises type by neighbourhoods.

Noted that there might be missing values as there might be no cases from certain premise type at specific neighbourhood (for instance no bicycle theft cases in Apartments at Bayview Woods-Steeles). Hence, we replace the missing values with 0.

```{r, message=FALSE}
# transform data
first_phase_bike_theft_clustering = bike_theft %>%
  group_by(NeighbourhoodName, Premises_Type) %>%
  summarise(count = n()) %>%
  pivot_wider(NeighbourhoodName, names_from = Premises_Type, values_from =count)

first_phase_bike_theft_clustering = data.frame(first_phase_bike_theft_clustering)

# replace missing values with 0
first_phase_bike_theft_clustering[is.na(first_phase_bike_theft_clustering)] <- 0

first_phase_bike_theft_clustering %>%
  arrange(desc(Outside))
```

After we transform the data, clustering analysis requires us to let the data in numerical standardize values only. To solve this, we let Neighbourhood names as row names instead of being part of data frame and standardize the values.

```{r}
# Transform the data into ready for K Means Clustering analysis

first_phase_k_means_bike_theft <- first_phase_bike_theft_clustering %>%
  select(-NeighbourhoodName)
first_phase_k_means_bike_theft <- scale(first_phase_k_means_bike_theft)
rownames(first_phase_k_means_bike_theft) = first_phase_bike_theft_clustering$NeighbourhoodName
```

### Determine K value using Elbow and Sillhoutte Method

Using Elbow Method, we can draw conclusion that *optimal K value is 2* because this is the point where total within sum of square drops a lot.

Using Silhouette Method, we can draw conclusion that *optimal K value is 2* because this is the point where average silhouette width is higher.

```{r}
elbow_graph <- fviz_nbclust(first_phase_k_means_bike_theft, kmeans, method = 'wss') +
  labs(title = "Optimal Number of Clusters - Elbow Method")+
  theme(
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 8),
    plot.title = element_text(size = 10, face = "bold", hjust = 0.9)
  )

silhouette_graph <- fviz_nbclust(first_phase_k_means_bike_theft, kmeans, method = 'silhouette') +
  labs(title = "Optimal Number of Clusters - Silhouette Method")+
  theme(
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 8),
    plot.title = element_text(size = 10, face = "bold", hjust = 0.9)
  )

ggarrange(elbow_graph,
          NULL,
          silhouette_graph,
          nrow = 1,
          ncol = 3,
          widths = c(1, 0.1, 1)
          )
```

### First Stage K-Means Clustering Result

The finding from following cluster plot is aligned with our previous argument from exploratory data analysis, which we can classify neighbourhoods based on bicycle theft premises type.

The top 10 neighbourhoods that contribute to 47% bicycle theft cases can classify into a single group where bicycle theft mostly happened outside building premises.

Meanwhile, the rest 130 neighbourhoods will be classified into another group where the number of bicycle theft cases are mostly shared in outside building premises, apartments and houses.

```{r}
# perform first round k-means clustering
first_phase_k_means_clustering = kmeans(first_phase_k_means_bike_theft, centers = 2, nstart = 50)

# visualise the result
fviz_cluster(first_phase_k_means_clustering, 
             data = first_phase_k_means_bike_theft, 
             labelsize = 6, 
             main = "Cluster Plot of Neighbourhood in Toronto", 
             ggtheme=theme_bw())
```

```{r}
#mapping cluster group number to Neighbourhood Name
bike_theft_by_neighbourhood %>%
  left_join(tibble::rownames_to_column(data.frame(first_phase_k_means_clustering$cluster)), by=c("NeighbourhoodName" = "rowname"))
```


## Exploring Cases by Bicycle Type

```{r, message=FALSE}
# bar chart of number of cases across neighbourhoods by bicycle type
bicycle_type_overall <- bike_theft %>%
  group_by(Bike_Type) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x= reorder(Bike_Type, -n), y= n)) + 
  geom_bar(stat='identity', fill = "#2B60DE") +
  geom_text(aes(label=n), position=position_dodge(width=0.5), vjust= -0.5, size = 3) +
  xlab("Bicycle Type") + 
  ylab("Number of Cases") +
  ggtitle("Overall Neighbourhoods") +
  theme_bw()+
  ylim(0, 10000)+
  theme(
    legend.title = element_blank(),
    plot.title = element_text(size = 10, hjust = -0.095, face ="bold"),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 9)
  )

# boxplot of number of cases in top 10 neighbourhoods by bicycle type
bicycle_type_top_10 <- bike_theft %>%
  filter(NeighbourhoodName %in% c(unique(top_10_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, Bike_Type) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(Bike_Type, levels = c("MT", "RG", "OT", "RC", "EL", "TO", "BM", "SC", "FO", "TR", "TA", "RE", "UN")), y = count)) + 
  geom_boxplot(fill = "#FFDB6D", color = "#C4961A")+
  xlab("Bicycle Type") +
  ylab("Number of Cases") +
  ggtitle("Top 10") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# boxplot of number of cases in next 32 neighbourhoods by bicycle type
bicycle_type_next_32 <- bike_theft %>%
  filter(NeighbourhoodName %in% c(unique(next_top_32_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, Bike_Type) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(Bike_Type, levels = c("MT", "RG", "OT", "RC", "EL", "TO", "BM", "SC", "FO", "TR", "TA", "RE", "UN")), y = count)) + 
  geom_boxplot(color = "#306754", fill = "#50C878") +
  xlab("Bicycle Type") +
  #ylab("Number of Cases") +
  ggtitle("Next Top 32") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# boxplot of number of cases in rest 98 neighbourhoods by bicycle type
bicycle_type_others <- bike_theft %>%
  filter(NeighbourhoodName %in% c(unique(other_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, Bike_Type) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(Bike_Type, levels = c("MT", "RG", "OT", "RC", "EL", "TO", "BM", "SC", "FO", "TR", "TA", "RE", "UN")), y = count)) + 
  geom_boxplot(color = "#F70D1A", fill = "#FF69B4") +
  xlab("Bicycle Type") +
  #ylab("Number of Cases") +
  ggtitle("Rest of Neighbourhoods") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# combine altogether in a plot
annotate_figure(ggarrange(bicycle_type_overall,
                          ggarrange(bicycle_type_top_10,
                                    bicycle_type_next_32, 
                                    bicycle_type_others, 
                                    nrow = 1, 
                                    ncol = 3),
                          nrow = 2,
                          ncol = 1),
                top = text_grob(expression(atop(bold("Number of Bicycle Theft Cases in Toronto by Bicycle Type (2014 - 2020)"), scriptstyle("Top 10 Neighbourhoods represent 47% of total cases, while Next Top 32 and The Rest contribute to 33% and 20% respectively")))))


```
<br>
Mountain Bike and Regular Bike become two most reported stolen bicycles in Toronto during 2014 - 2020, and this also show across neighbourhood groups.

## Exploring Cases by Reported Day

```{r, message=FALSE}
# bar chart of number of cases across neighbourhoods by reported day
reported_day_overall <- bike_theft %>%
  mutate(weekday_name = weekdays(Report_Date)) %>%
  group_by(weekday_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(weekday_name, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")), y = count)) + 
  geom_bar(stat='identity', fill = "#2B60DE") +
  geom_text(aes(label=count), position=position_dodge(width=0.5), vjust= -0.5, size = 3) +
  ylim(0,5000) +
  xlab("Day of Week") +
  ylab("Number of Cases") +
  ggtitle("Overall Neighbourhoods") +
  theme_bw()+
  theme(
    legend.title = element_blank(),
    plot.title = element_text(size = 10, hjust = -0.095, face ="bold"),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 9)
  )

# boxplot of number of cases in top 10 neighbourhoods by reported day
reported_day_top_10 <- bike_theft %>%
  mutate(weekday_name = weekdays(Report_Date)) %>%
  filter(NeighbourhoodName %in% c(unique(top_10_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, weekday_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(weekday_name, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")), y = count)) + 
  geom_boxplot(fill = "#FFDB6D", color = "#C4961A") +
  xlab("Day of Week") +
  ylab("Number of Cases") +
  ggtitle("Top 10") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# boxplot of number of cases in next 32 neighbourhoods by reported day
reported_day_next_32 <- bike_theft %>%
  mutate(weekday_name = weekdays(Report_Date)) %>%
  filter(NeighbourhoodName %in% c(unique(next_top_32_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, weekday_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(weekday_name, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")), y = count)) +
  geom_boxplot(color = "#306754", fill = "#50C878") +
  xlab("Day of Week") +
  #ylab("Number of Cases") +
  ggtitle("Next Top 32") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# boxplot of number of cases in rest 98 neighbourhoods by reported day
reported_day_others <- bike_theft %>%
  mutate(weekday_name = weekdays(Report_Date)) %>%
  filter(NeighbourhoodName %in% c(unique(other_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, weekday_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(weekday_name, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")), y = count)) +
  geom_boxplot(color = "#F70D1A", fill = "#FF69B4") +
  xlab("Day of Week") +
  #ylab("Number of Cases") +
  ggtitle("Rest of Neighbourhoods") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

#combine altogether in a plot
annotate_figure(ggarrange(reported_day_overall,
                          ggarrange(reported_day_top_10,
                                    reported_day_next_32, 
                                    reported_day_others, 
                                    nrow = 1, 
                                    ncol = 3),
                          nrow = 2,
                          ncol = 1),
                top = text_grob(expression(atop(bold("Number of Bicycle Theft Cases in Toronto by Reported Day (2014 - 2020)"), scriptstyle("Top 10 Neighbourhoods represent 47% of total cases, while Next Top 32 and The Rest contribute to 33% and 20% respectively")))))

```
<br>
From the graph above, more theft cases were happened during weekdays and this is applied across neighbourhood groups. 

## Exploring Cases by Reported Month

```{r, message=FALSE}
# bar chart of number of cases across neighbourhoods by reported month
reported_month_overall <- bike_theft %>%
  mutate(month_name = month.name[month(Report_Date)]) %>%
  group_by(month_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(month_name, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")), y = count)) + 
  geom_bar(stat='identity', fill = "#2B60DE") +
  geom_text(aes(label=count), position=position_dodge(width=0.5), vjust= -0.5, size = 3) +
  ylim(0,5000) +
  xlab("Month") +
  ylab("Number of Cases") +
  ggtitle("Overall Neighbourhoods") +
  theme_bw()+
  theme(
    legend.title = element_blank(),
    plot.title = element_text(size = 10, hjust = -0.095, face ="bold"),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 9),
    axis.text.x = element_text(size = 8)
  )

# boxplot of number of cases in top 10 neighbourhoods by reported day
reported_month_top_10 <- bike_theft %>%
  mutate(month_name = month.name[month(Report_Date)])  %>%
  filter(NeighbourhoodName %in% c(unique(top_10_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, month_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(month_name, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")), y = count)) + 
  geom_boxplot(fill = "#FFDB6D", color = "#C4961A") +
  xlab("Month") +
  ylab("Number of Cases") +
  ggtitle("Top 10") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# boxplot of number of cases in next 32 neighbourhoods by reported day
reported_month_next_32 <- bike_theft %>%
  mutate(month_name = month.name[month(Report_Date)]) %>%
  filter(NeighbourhoodName %in% c(unique(next_top_32_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, month_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(month_name, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")), y = count)) + 
  geom_boxplot(color = "#306754", fill = "#50C878") +
  xlab("Month") +
  #ylab("Number of Cases") +
  ggtitle("Next Top 32") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

# boxplot of number of cases in rest 98 neighbourhoods by reported day
reported_month_others <- bike_theft %>%
  mutate(month_name = month.name[month(Report_Date)]) %>%
  filter(NeighbourhoodName %in% c(unique(other_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, month_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(month_name, levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")), y = count)) + 
  geom_boxplot(color = "#F70D1A", fill = "#FF69B4") +
  xlab("Month") +
  #ylab("Number of Cases") +
  ggtitle("Rest of Neighbourhoods") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(angle = 90, size = 7, hjust = 1)
  )

#combine altogether in a plot
annotate_figure(ggarrange(reported_month_overall,
                          ggarrange(reported_month_top_10,
                                    reported_month_next_32, 
                                    reported_month_others, 
                                    nrow = 1, 
                                    ncol = 3),
                          nrow = 2,
                          ncol = 1),
                top = text_grob(expression(atop(bold("Number of Bicycle Theft Cases in Toronto by Reported Month (2014 - 2020)"), scriptstyle("Top 10 Neighbourhoods represent 47% of total cases, while Next Top 32 and The Rest contribute to 33% and 20% respectively")))))

```
<br>
We also observed seasonality in number of bicycle theft cases, where the cases start to hike up during spring, reached its highest peak during summer and plummeted from fall season. This pattern is also reflected across neighbourhood groups. 


## Exploring Cases by Reported Year

### Identify trend by neighbourhood groups

```{r, message=FALSE}
# bar chart of number of cases across neighbourhoods by reported year
reported_year_overall <- bike_theft %>%
  mutate(year_name = year(Report_Date)) %>%
  group_by(year_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(year_name), y = count)) + 
  geom_bar(stat='identity', fill = "#2B60DE") +
  geom_text(aes(label=count), position=position_dodge(width=0.5), vjust= -0.5, size = 3) +
  ylim(0,5000) +
  xlab("Year") +
  ylab("Number of Cases") +
  ggtitle("Overall Neighbourhoods") +
  theme_bw()+
  theme(
    legend.title = element_blank(),
    plot.title = element_text(size = 10, hjust = -0.095, face ="bold"),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 9)
  )

# boxplot of number of cases in top 10 neighbourhoods by reported day
reported_year_top_10 <- bike_theft %>%
  mutate(year_name = year(Report_Date)) %>%
  filter(NeighbourhoodName %in% c(unique(top_10_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, year_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(year_name), y = count)) + 
  geom_boxplot(fill = "#FFDB6D", color = "#C4961A") + 
  xlab("Year") +
  ylab("Number of Cases") +
  ggtitle("Top 10") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(size = 7, hjust = 0.5)
  )

# boxplot of number of cases in next 32 neighbourhoods by reported day
reported_year_next_32 <- bike_theft %>%
  mutate(year_name = year(Report_Date)) %>%
  filter(NeighbourhoodName %in% c(unique(next_top_32_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, year_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(year_name), y = count)) +
  geom_boxplot(color = "#306754", fill = "#50C878") +
  xlab("Year") +
  #ylab("Number of Cases") +
  ggtitle("Next Top 32") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(size = 7, hjust = 0.5)
  )

# boxplot of number of cases in rest 98 neighbourhoods by reported day
reported_year_others <- bike_theft %>%
  mutate(year_name = year(Report_Date)) %>%
  filter(NeighbourhoodName %in% c(unique(other_neighbourhood$NeighbourhoodName))) %>%
  group_by(NeighbourhoodName, year_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x=factor(year_name), y = count)) +
  geom_boxplot(color = "#F70D1A", fill = "#FF69B4") +
  xlab("Year") +
  #ylab("Number of Cases") +
  ggtitle("Rest of Neighbourhoods") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 10, face ="bold"),
    axis.title.y = element_blank(),
    axis.title = element_text(size = 8),
    axis.text.x = element_text(size = 7, hjust = 0.5)
  )

#combine altogether in a plot
annotate_figure(ggarrange(reported_year_overall,
                          ggarrange(reported_year_top_10,
                                    reported_year_next_32, 
                                    reported_year_others, 
                                    nrow = 1, 
                                    ncol = 3),
                          nrow = 2,
                          ncol = 1),
                top = text_grob(expression(atop(bold("Number of Bicycle Theft Cases in Toronto by Reported Year (2014 - 2020)"), scriptstyle("Top 10 Neighbourhoods represent 47% of total cases, while Next Top 32 and The Rest contribute to 33% and 20% respectively")))))

```
<br>
It is noted that in Next Top 32, the median value number of cases was increasing in 2020 while the other groups were stagnant.

### Identify which neighborhoods that contribute the most to the increase of bicycle theft cases in 2020

The graph below depicts top contributors to the incremental trend number of cases in 2020.

```{r, message=FALSE, fig.width = 10}
# annotate neighbourhood groupings
neighbourhood_grouping <- bike_theft %>%
  mutate(grouping = ifelse(NeighbourhoodName %in% c(unique(top_10_neighbourhood$NeighbourhoodName)),
                           "Top 10",
                           ifelse(NeighbourhoodName %in% c(unique(next_top_32_neighbourhood$NeighbourhoodName)),
                                  "Next Top 32", "Rest of Neighbourhoods")),
         year_name = year(Report_Date))

# create graph year-on-year trend by neighbourhood group
trend_by_neighbourhood_grouping_graph <- neighbourhood_grouping %>%
  group_by(year_name, grouping) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = year_name, y = count, fill = factor(grouping, levels = c("Top 10", "Next Top 32", "Rest of Neighbourhoods")), label = count)) +
  geom_bar(stat = "identity") +
  geom_text(size = 3, position = position_stack(vjust = 0.5))+
  ylim(0,5000) +
  xlab("Year") +
  ylab("Number of Cases") +
  ggtitle("Y-o-Y bicycle theft cases by neighbourhood groups")+
  guides(fill=guide_legend(title="Neighbourhood Groupings")) +
  theme_bw()+
  scale_fill_manual(values = c("#FFDB6D", "#50C878", "#FF69B4"))+
  scale_x_continuous(breaks = seq(2014, 2020, 1)) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 7),
    plot.title = element_text(size = 10, hjust = 0, face ="bold"),
    axis.title = element_text(size = 9),
    axis.text = element_text(size = 9)
  )

# calculate change from 2019 to 2020, and take only top 10
top_10_incremental_2020 <- neighbourhood_grouping %>%
  mutate(year_name = paste("year_", year(Report_Date), sep="")) %>%
  filter(year_name %in% c("year_2019", "year_2020")) %>%
  group_by(NeighbourhoodName, year_name, grouping) %>%
  summarise(count = n()) %>%
  pivot_wider(c(NeighbourhoodName, grouping), names_from = year_name, values_from =count) %>%
  replace_na(list(year_2019 = 0, year_2020 = 0)) %>%
  mutate(change = year_2020 - year_2019) %>%
  arrange(desc(change)) %>%
  head(10)

# create graph to show the changes of top 10 contributors
top_10_incremental_2020_graph <- top_10_incremental_2020 %>%
  ggplot(aes(x = reorder(NeighbourhoodName, change), 
             y = change, 
             fill = grouping,
             label = change)) +
  geom_bar(stat = "identity") +
  geom_text(size = 3, position = position_stack(vjust = 0.5))+
  ylim(0,80) +
  scale_fill_manual(values = c("Top 10" = "#FFDB6D", "Next Top 32" = "#50C878", "Rest of Neighbourhoods" = "#FF69B4"))+
  xlab("Neighbourhood") +
  ylab("Increase in Number of Cases from 2019") +
  ggtitle("Top 10 contributors to incremental cases in 2020")+
  theme_bw()+
  theme(
    legend.title = element_text(size = 8, face = "bold"),
    legend.text = element_text(size = 8),
    plot.title = element_text(size = 10, face ="bold", hjust = 10),
    axis.title = element_text(size = 7, face = "bold"),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none"
  ) + 
  coord_flip()
 
# combine graphs altogether 
ggarrange(trend_by_neighbourhood_grouping_graph,
          NULL,
          top_10_incremental_2020_graph,
          nrow = 1,
          widths = c(1, 0.1, 1)
          )
  
```
<br>
From the chart above, we can see Next Top 32 contribute the incremental number of cases in 2020 by 252, followed by rest of neigbourhoods (221).

Take top 10 contributors, we can see 6 out of 10 were coming from Next Top 32.

### Identify which premise type that contribute the most to the increase of bicycle theft cases in 2020

```{r, message=FALSE, fig.width = 10}
# stacked barchart showing year-on-year trend by premise type
trend_by_premise_graph <- neighbourhood_grouping %>%
  group_by(year_name, Premises_Type) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = year_name, 
             y = count, 
             fill = factor(Premises_Type, 
                           levels = c("Outside", "Apartment", "House", "Commercial", "Other", "Educational", "Transit")), 
             label = count)) +
  geom_bar(stat = "identity") +
  geom_text(size = 3, position = position_stack(vjust = 0.5))+
  ylim(0,5000) +
  xlab("Year") +
  ylab("Number of Cases") +
  ggtitle("Y-o-Y Bicycle Theft Cases by Premise Type")+
  guides(fill=guide_legend(title="Premise Type")) +
  theme_bw()+
  scale_x_continuous(breaks = seq(2014, 2020, 1)) +
  scale_fill_manual(values = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"))+
  theme(
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 7),
    plot.title = element_text(size = 10, face ="bold"),
    axis.title = element_text(size = 7, face = "bold"),
    axis.text = element_text(size = 7)
  )

# calculate changes on number of cases from 2019 to 2020 by premise type, each neighbourhood
top_10_incremental_2020_by_premise <- neighbourhood_grouping %>%
  mutate(year_name = year(Report_Date)) %>%
  filter(year_name %in% c(2019, 2020)) %>%
  group_by(NeighbourhoodName, year_name, Premises_Type) %>%
  summarise(count = n()) %>%
  pivot_wider(NeighbourhoodName, names_from = c(Premises_Type, year_name), values_from =count) %>%
  replace_na(list(Apartment_2019 = 0, Apartment_2020 = 0,
                  House_2019 = 0, House_2020 = 0,
                  Outside_2019 = 0, Outside_2020 = 0,
                  Transit_2019 = 0, Transit_2020 = 0,
                  Educational_2019 = 0, Educational_2020 = 0,
                  Commercial_2019 = 0, Commercial_2020 = 0,
                  Other_2019 = 0, Other_2020 = 0)) %>%
  mutate(
    Apartment = Apartment_2020 - Apartment_2019,
    House = House_2020 - House_2019,
    Outside = Outside_2020 - Outside_2019,
    Transit = Transit_2020 - Transit_2019,
    Educational = Educational_2020 - Educational_2019,
    Commercial = Commercial_2020 - Commercial_2019,
    Other = Other_2020 - Other_2019,
    Overall = Apartment + House + Outside + Transit + Educational + Commercial + Other
  ) %>%
  arrange(desc(Overall)) %>%
  head(10) %>%
  select(-c(Apartment_2019, Apartment_2020,
            House_2019, House_2020,
            Outside_2019, Outside_2020,
            Transit_2019, Transit_2020,
            Educational_2019, Educational_2020,
            Commercial_2019, Commercial_2020,
            Other_2019, Other_2020,
            Overall)) %>%
  pivot_longer(!NeighbourhoodName, names_to = "Premises_Type", values_to = "Count")

# create bar chart for top 10 contributors by premise type
top_10_incremental_2020_by_premise_graph <- top_10_incremental_2020_by_premise %>%
  ggplot(aes(x = factor(NeighbourhoodName, 
                        levels= rev(c(unique(top_10_incremental_2020_by_premise$NeighbourhoodName)))), 
             y = Count, 
             fill = factor(Premises_Type, 
                           levels = c("Outside", "Apartment", "House", "Commercial", "Other", "Educational", "Transit")),
             label = Count
             )
         ) +
  theme_bw()+
  geom_bar(stat="identity") +
  geom_text(size = 2, position = position_stack(vjust = 0.5))+
  guides(fill=guide_legend(title="Premise Type")) +
  xlab("Neighbourhood") +
  ylab("Increase in Number of Cases from 2019") +
  ggtitle("Top 10 Contributors by Premise Type")+
  scale_fill_manual(values = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"))+
  theme(
    legend.title = element_text(size = 8, face = "bold"),
    legend.text = element_text(size = 8),
    plot.title = element_text(size = 10, face ="bold"),
    axis.title = element_text(size = 7, face = "bold"),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    legend.position = "none"
  ) + 
  coord_flip()

# combine graphs altogether
ggarrange(trend_by_premise_graph,
          NULL,
          top_10_incremental_2020_by_premise_graph,
          nrow = 1,
          widths = c(1, 0.1, 1)
          )
  
```
<br>
The number of bicycle theft cases at apartments shows a significant contribution to the increase in the number of cases in 2020. Further, this figure also contributes most to the incremental number of cases in Mount Pleasant West (49 cases), Willowdale East (28), and Bayview Village (22).
<br>

# Conclusion

From the analysis, we can draw insights as follow.

-   10 neighbourhoods are contributed to around 47% of overall reported bicycle theft cases in Toronto from 2014 - 2020. Majority of cases from this neighbourhood group happened when the bicycle is parked outside building premises.

-   Apartments in Mount Pleasant View, Willowdale East, and Bayview Village; as these provide significant contribution to the incremental number of bicycle theft cases in 2020.


